Sketch Recognition using Particle Swarm Optimizations

### 2009-06-06 paper2/Paper2.tex
after all editin gis done

Yu \cite{meanshift10} introduced a \textit{feature area} for each primitive and then computed the segmentation error for different types of primitives based on the computed \textit{feature area}. His system achieved good accuracy in simple shapes (square, ellipse,...etc) but did not perform well in complex shapes. A hybrid algorithm was introduced in \cite{earlyprocess} where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set. However, this system is limited to recognizing only specific simple geometric shapes with a set of low level recognizers. Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from input stroke.

@@

\begin{table}
\begin{center}
\scalebox{0.9}{
 \begin{tabular}{|p{2cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
 \hline
 & Hs-DS &  EL-DS & LD-DS \\ \hline
No of samples  &  7791 & 2764 & 1859 \\ \hline
No of categories  & 13 & 15 & 8 \\ \hline
Avg. Samples per Category& 600 & 184 & 232 \\\hline 
No of users  & 16 & 7 & 7 \\  \hline
Balanced & Yes&No & No \\ \hline
Splits  & 5 random splits  & 5 random splits  & 5 random splits  \\ \hline
Examples & Figure \ref{fig:HsSet}  & Figure \ref{fig:ELsymbolSet}  & Figure \ref{fig:LogicsymbolSet}   \\ \hline
\end{tabular}
}
\caption[Datasets Comparisons]{Table compares between datasets used in experiments in terms of number of samples, number of samples per category, etc...} 
\label{tab:datasets}

@@

 \begin{figure}
	\centering		
	 \includegraphics[scale=0.3]{images/swarmAgentTest.jpg}
	 	\caption{Segmentation Error Vs. Number of particles} The error decrease as the number of particle increase.  
	 	\label{fig:swarmtesting}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

\begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{images/swarmParamterC.jpg}
	 	\caption{Segmentation Error Vs. PSO parameters.} The behaviour of segmentation error versus c1. 
	 	\label{fig:swarmtesting2}
	%\caption{Experiments results :} a)    b) c)

\end{figure} 
\subsection{Algorithms comparisons}
\label{sec:AlgExp}
We performed several experiments to evaluate the presented recognition system. Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms. Figure \ref{fig:test1} and \ref{fig:testLD} shows the accuracy achieved by each algorithm. The two swarm algorithms were tested with and without the ellipse fitting module. Combining the ellipse detection module improves the performance. The result shows that combining both (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm. %show that combining algoirthmsachieve higher accuracy than any single algoirthm. 
%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.

 \begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{images/testAlg.jpg}
	 	\caption{Algorithm comparison on HS-DS} The recognition rate of different algorithms. 
	 	\label{fig:test1}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

\begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{images/LDAlg.jpg}
	 	\caption{Algorithm comparison on LD-DS} The recognition rate of different algorithms. 
	 	\label{fig:testLD}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

\begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{images/AlgEL.jpg}
	 	\caption{Algorithm comparison on EL-DS} The recognition rate of different algorithms. 
	 	\label{fig:testEL}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

\subsection{Shape Complexity Analysis}
\label{sec:ShapeComplexityExperiments}

\begin{figure*}
	\centering
		\includegraphics[scale=0.5]{images/testsym.jpg}
	\caption{Symbols Comparison on HS-DS} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:test2}
\end{figure*}

@@

		\includegraphics[scale=0.5]{images/LDsymbols.jpg}
	\caption{Symbols Comparison on LD-DS symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:LDtest2}

@@

		\includegraphics[scale=0.5]{images/featuresdifference.jpg}
	\caption{Feature Comparison on Hs-DS accuracy} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeaturesAll}

 \begin{figure*}
	\centering
		\includegraphics[scale=0.34]{images/featELc.jpg}
	\caption{Feature Comparison on EL-DS} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:ELtestFeaturesAll}
\end{figure*}  
 We could not test the result of the segmentation algorithm alone due to the fact that the correct segmentation is hightly qualitative. It is only based on what the user intends to draw while sketching the symbol. 
 The main features affected by the segmentation stage are the geometric features computed in \textbf{FS1}. Hence those are the features we used to estimate the efficiency of the segmentation algorithm. The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} . The figure shows that \textsl{AlgS2} achieve better results.  %effect of the segmentation algorithm 
 %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block. The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}). Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features. 
\begin{figure}
	\centering
		\includegraphics[scale=0.4]{images/featureFS1only.jpg}
	\caption{The effect of segmentation on Hs-DS accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeatonly}
\end{figure}

@@

	\caption{Segmentation outputs}  The output segmentation of different input strokes. %The figure shows sample of the output of the segmentation phase. Figure a) shows the stroke after the user draws it. The red points are points labeled as \textit{possible dominate points}. Figure b) shows the symbol after being segmented. Figure c) shows the segmentation of different and overlapped strokes. Notice the system mark the center of the ellipse that was identified.
	\label{fig:SampleSeg}
\end{figure}

@@

This paper presented a new approach to sketch recognition using PSO. The system uses both sthe peed and curvature information as input to the \textit{DPSO}. It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The tradeoff between accuracy achieved and time complexity can be further investigated to achieve better results. The final recognition is based on both global shape and hierarchical stroke properties. Adding stroke and segment geometrical features improved the final recognition. Global shape and moment properties were used to prevent the need to use graph matching which is computationally expensive \cite{SketchRead2007}. The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%. The system does not depend on low level recognizer but rather on a set of high level features. This makes the system easily expandable to more and more symbols.

>>>>

Yu \cite{meanshift10} introduced a \textit{feature area} for each primitive and then computed the segmentation error for different types of primitives based on the computed \textit{feature area}. His system achieved good accuracy in simple shapes (square, ellipse,...etc) but did not perform well in complex shapes. A hybrid algorithm was introduced in \cite{earlyprocess}\footnote{Alg3 used in Section \ref{sec:Experiments} } where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set. However, this system is limited to recognizing only specific simple geometric shapes with a set of low level recognizers. Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from input stroke.

@@

\begin{table}
\begin{center}
\scalebox{0.8}{
 \begin{tabular}{|p{3cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
 \hline
 & Hs-DS &  EL-DS & LD-DS \\ \hline
No of samples  &  7791 & 2764 & 1859 \\ \hline
No of categories  & 13 & 15 & 8 \\ \hline
Avg. Samples per Category& 600 & 184 & 232 \\\hline 
No of users  & 16 & 7 & 7 \\  \hline
Balanced & Yes&No & No \\ \hline
Splits  & 5 random splits  & 5 random splits  & 5 random splits  \\ \hline
Examples & Figure \ref{fig:HsSet}  & Figure \ref{fig:ELsymbolSet}  & Figure \ref{fig:LogicsymbolSet}   \\ \hline
\end{tabular}
}
\caption[Datasets Comparisons]{Table compares between datasets used in experiments in terms of number of samples, number of samples per category, etc...} 
\label{tab:datasets}

@@

 \begin{figure}
	\centering		
	 	\subfigure[Segmentation Error Vs. Number of particles]{	\label{fig:swarmtesting} \includegraphics[scale=0.3]{images/swarmAgentTest.jpg}}
	 %	\caption{}

	%\caption{Experiments results :} a)    b) c) 
	%\end{figure}

%\begin{figure}
	%\centering		
		\subfigure[Segmentation Error Vs. PSO parameters.]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.4]{images/swarmParamterC.jpg}}
	 	\caption{}a) The error decrease as the number of particle increase.   b) The behaviour of segmentation error versus c1.

	%\Segmentation Error Vs. PSO parameters.caption{Experiments results :} a)    b) c)

\end{figure} 
\subsection{Algorithms comparisons}
\label{sec:AlgExp}
We performed several experiments to evaluate the presented recognition system. Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms. Figure \ref{fig:test1} and \ref{fig:testLD} shows the accuracy achieved by each algorithm. The two swarm algorithms were tested with and without the ellipse fitting module. Combining the ellipse detection module improves the performance. The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1 in all datasets.  The results also shows that combineing  (\textsl{AlgS1} and \textsl{AlgS2} using minimum error) outperforme AlgS2 in Hs-Ds only. This behaviour is explained by noticing that features are computed based on the choosen segmentation. This means that different segmentation for a symbol generates different set of features used for recognition. This behaviour affect Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure symbols in both datasets than than of Hs-Ds.  %   can be explained by the simplicity of the symbols in Hs-DS % and LD-Ds unlike EL-DS where AlgS2 is the best algorithm. This behaviour is explained by the structure of the symbols in each dataset. El-DS consist of highly topological and structure symbols that are usaually drawn using multiple small strokes. This leads to    %   out perform any single algorithm in HS-DS and LD-DS.  %show that combining algoirthmsachieve higher accuracy than any single algoirthm. 
%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.

 \begin{figure}[h]
	\centering		
	\subfigure[Algorithm comparison on HS-DS]{\label{fig:test1} \includegraphics[scale=0.4]{images/testAlg.jpg}}
	 %	\caption{} The recognition rate of different algorithms.

	%\caption{Experiments results :} a)    b) c) 
%\end{figure}

%\begin{figure}[h]
%	\centering		
		\subfigure[Algorithm comparison on LD-DS]{ 	 	\label{fig:testLD} \includegraphics[scale=0.4]{images/LDAlg.jpg}}
	% 	\caption{} The recognition rate of different algorithms.

	%\caption{Experiments results :} a)    b) c) 
%\end{figure}

%\begin{figure}[h]
%	\centering		
		\subfigure[Algorithm comparison on EL-DS]{	 	\label{fig:testEL}  \includegraphics[scale=0.4]{images/AlgE.jpg}}
	 	\caption{ The recognition rate of different algorithms. }

	%\caption{Experiments results :} a)    b) c) 
\end{figure}

\subsection{Shape Complexity Analysis}
\label{sec:ShapeComplexityExperiments}

@@

		\subfigure[Symbols Comparison on HS-DS]{\label{fig:test2}
		\includegraphics[scale=0.35]{images/testsym.jpg}}
	%\caption{} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
%\end{figure*}  
%\begin{figure*}
\subfigure[Symbols Comparison on LD-DS symbols]{\label{fig:LDtest2}
		\includegraphics[scale=0.5]{images/DLSymbols.jpg}}
	%\caption{}.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\caption {Symbol Comparison} The accuarcy of achieved by each algorithm in each symbol. a) Low accuracy level appears in triangle, heart, and moon symbols. b) Low accuracy level appears only in symbols that are visually confusing (i.e Nor, OR, Nand gates). 

@@

	\subfigure[Feature Comparison on Hs-DS]{	\label{fig:testFeaturesAll} \includegraphics[scale=0.5]{images/featuresdifference.jpg}}
    %The graph shows the recognition rate of each symbol using different algorithms.

%\end{figure*}

 %\begin{figure*}
	%\centering

	\subfigure[Feature Comparison on EL-DS]{	\label{fig:ELtestFeaturesAll}	\includegraphics[scale=0.34]{images/FeatEl.jpg}}
	\caption{Different Features } The effect of different feature set on accuracy.  FS4 and FS3 give good results in a) Hs-DS and b) EL-Ds. The results is improved by combining FS1 with FS2 and FS1 with FS4.    %The graph shows the recognition rate of each symbol using different algorithms.

\end{figure*}  
 We could not test the result of the segmentation algorithm alone due to the fact that the correct segmentation is hightly qualitative. It is only based on what the user intends to draw while sketching the symbol. 
 The main features affected by the segmentation stage are the geometric features computed in \textbf{FS1}. Hence those are the features we used to estimate the efficiency of the segmentation algorithm. The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly}. The figure shows that \textsl{AlgS2} achieve better results.  %effect of the segmentation algorithm 
 %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block. The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}). Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features. 
\begin{figure}
	\centering
		\includegraphics[scale=0.3]{images/OnlyF1EL.jpg}
	\caption{The effect of segmentation on Hs-DS accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeatonly}
\end{figure}

@@

	\caption{Segmentation outputs}  The output segmentation of different input strokes. a) Sample of single input stroke b) The single stroke after segmentation. c) An example of multiple stroke and their segmentation. %The figure shows sample of the output of the segmentation phase. Figure a) shows the stroke after the user draws it. The red points are points labeled as \textit{possible dominate points}. Figure b) shows the symbol after being segmented. Figure c) shows the segmentation of different and overlapped strokes. Notice the system mark the center of the ellipse that was identified.
	\label{fig:SampleSeg}
\end{figure}

@@

This paper presented a new approach to sketch recognition using DPSO. The system uses multiple premilary information i.e., speed and curvature information, as input to the \textit{DPSO}. It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The tradeoff between accuracy achieved and time complexity can be further investigated to achieve better results. The final recognition is based on both global shape and hierarchical stroke properties. Adding stroke and segment geometrical features improved the final recognition. Global shape and moment properties were used to prevent the need to use graph matching which is computationally expensive \cite{SketchRead2007}. The experiments evaluated two\textit{DPSO} different segmenatation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds, EL-Ds and LD-Ds). Results show that AlgS2 achieves better segmentation  and final accuracy than Alg3 \cite{earlyprocess} on the datasets used. The results proved that the system is easily expandable to more and more symbols as it dose not depend on low level primitive recogniziers but on high level symbol and geometrical featuers.  The system achieved an overall improvement of more than 2\% over Alg3 in all datasets.  %The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%. The system does not depend on low level recognizer but rather on a set of high level features. This makes the system easily expandable to more and more symbols.
